---
title: "POL 309: Political Psychology"
format:
    html: default
    pdf: 
      documentclass: article
      fontsize: 12pt
      colorlinks: true
      citecolor: DarkViolet
      linkcolor: DarkViolet
      urlcolor: DarkViolet
      geometry: margin=1in
      output-file: "POL309_syllabus"
---

```{r setup, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(tibble)
library(pander)
```

::: {.content-visible when-format="pdf"}

\begin{center}

\vspace{-5em}

\url{ {{< var course.url >}} } \vspace{-1ex}

{{< var course.time >}} {{< var course.days >}} \vspace{-1ex}

{{< var course.location >}} \vspace{-1ex}

{{< var course.semester >}}

\end{center}


## Instructor

{{< var instructor.name >}} \vspace{-1ex}

Email: [{{< var instructor.email >}}](mailto:{{< var instructor.email >}}) \vspace{-1ex}

Office: {{< var instructor.office >}} \vspace{-1ex}

Office hours: Thursdays 10 am - 12 pm

:::


::: {.content-visible when-format="html"}

::: {.grid .course-details .course-details-smaller}

::: {.g-col-6}
### Course details

- {{< fa calendar >}} &nbsp; {{< var course.days >}}
- {{< fa calendar-alt >}} &nbsp; {{< var course.dates >}}
- {{< fa clock >}} &nbsp; {{< var course.time >}}
- {{< fa location-dot >}} &nbsp; {{< var course.location >}}
:::

::: {.g-col-6}
### Instructor

- {{< fa user >}} &nbsp; [{{< var instructor.name >}}]({{< var instructor.url >}})
- {{< fa university >}} &nbsp; {{< var instructor.office >}}
- {{< fa envelope >}} &nbsp; [{{< var instructor.email >}}](mailto:{{< var instructor.email >}})
- {{< fa university >}} &nbsp; Office hours: Thursdays 10 am - 12 pm
:::

:::



# Syllabus

You can find a PDF version of the syllabus with the weekly schedule on the sidebar of this page.

:::


## Course Overview

Why do people vote the way they do? How do racial and gender stereotypes shape how voters evaluate candidates for elective office? How does the media influence political attitudes? Do anger and anxiety mobilize or demobilize voters? How can we answer these questions? This course examines these questions and the answers that political psychologists provide us by applying concepts from social and cognitive psychology to political phenomena. In addition to understanding how psychological processes shape how citizens engage in politics, students will gain familiarity with the tools political scientists and psychologists use to study these questions.

## Learning Goals

The American Psychological Association's Guidelines 3.0 for the Undergraduate Psychology Major consists of five goals, which are being used as the student learning objectives for Agnes Scott's Psychology majors. In this course, we will focus on goals 1, 2 and 4:

- Goal 1: Content Knowledge and Applications
- Goal 2: Scientific Inquiry and Critical Thinking
- Goal 4: Communication, Psychological Literacy, and Technology Skills

To work on these objectives:

- Students will read and interpret research articles from the primary literature in the topic of the course, and demonstrate comprehension through an analysis of these articles. They will then write a paper using APA (or APSA) writing style. Students will accurately summarize general ideas and conclusions from psychological sources for these assignments.

- Students will write a response paper to a week's readings where they will synthesize what political psychologists know and how they know it. They will also offer an application to current events and/or a critique of existing work.

- Students will learn about how psychological theories can be applied to a different field of study within the social sciences.

- Students will use their knowledge of culture, values, and biases learned in courses to reduce misunderstandings in communication during class discussions, assignments, and/or projects.



## Learning Outcomes

- Demonstrate an understanding of the main theories in political psychology
- Think critically about people’s political behavior and how to study it
- Identify lingering questions about political behavior and ways to answer them
- Apply concepts and theories from political psychology research to their political environment


## Work and Assessment

### Course Expectations

Students are expected to attend class and participate during class discussions. This course has both a lecture and a discussion component. The lectures will incorporate the readings assigned for that day/week. I expected you will have completed all readings prior to class and will come prepared with comments and questions. Readings can be accessed through the library resources or, when not available through the library, via Canvas.

- <ins>*A note of caution:*</ins> many, if not most, of the readings come from academic journals and books, mostly from political science but also from psychology. You will find that many of the articles are quite technical and they can be hard to follow even when one is familiar with the methodology. Do not despair. Make sure to understand, and wrestle with, the main ideas presented in a piece. Pay attention as well to how the authors support their arguments with data without getting bogged down by technical details.

### Course Workload

This is a 4-credit hour course, meaning that you are expected to spend 12 hours of work each week (including the 3 hours spent in the classroom). This roughly means that you will spend about 2 hours on readings for each class period, and 3 hours each week working on the different assignments (pace yourself!).


### Assignments and grades

Final course grades will be assigned as follows:

```{r grades, include=FALSE}
grading <- tribble(
  ~Grade, ~Range, ~Grade1, ~Range1,
  "A",  "93–100%", "C",  "73–76%",
  "A−", "90–92%", "C−", "70–72%",
  "B+", "87–89%", "D+", "67–69%",
  "B",  "83–86%", "D",  "63–66%",
  "B−", "80–82%", "D−", "60–62%",
  "C+", "77–79%", "F",  "< 60%"
)
```

<div class="centered-table">

```{r show-grades-table, echo=FALSE, results="asis"}
grading %>%
  pandoc.table(justify = "lclc",
               col.names = c("Grade", "Range", "Grade", "Range"))
```

<div>


Final grades will be calculated based on the following components:

**Participation (20%)**: All students are expected to be active participants in class discussions. Our learning experience will be enhanced if you come prepared and ready to share your comments, insights, and/or questions from the material covered in the lectures and readings with the rest of the class.

  - <ins>*Attendance policy:*</ins> Attendance is part of your participation grade, and while there is no separate grade for attendance it is a prerequisite for in-class participation. That said, to accommodate unforeseeable circumstances, you have two *unexcused* absences that will not affect your participation grade. Please let the instructor know about any excused absences you anticipate during the semester. Note that students who are absent, whether excused or not, are still expected to submit the weekly summary corresponding to that class period.


**Weekly summaries (20%)**: Each week for eight weeks, you will summarize one article from the week's readings (the reading you can summarize has a {{< fa pen >}} after it in the course schedule). There are 13 weeks in the semester where you can write a summary, so you can skip some weeks. It is up to you to decide which weeks to skip. The summary must be no longer than three-quarters of a page (single-spaced, 12-point font). These summaries must be submitted on Canvas before class time *on the day the reading was assigned*.


**Response paper (15%)**: Students will write a response paper (2 double-spaced pages, 12-point font) where they integrate the readings for a given week. In the response paper, you will highlight the connections between the readings, write about their applicability to current events, and offer a critique of the readings as a whole. You may choose to focus more on applicability to today or a critique of existing research. The response paper is due on the Friday of the week you choose to write on (11:59 pm deadline).


**Discussion leader (15%)**: Students will serve as discussion leaders for a portion of class time each week. At the beginning of the semester, students will choose one week where they will co-lead the discussion. You are expected to have done the readings and come up with discussion questions for the class. I encourage you to meet with me before your week to talk about the material. (P.S.: this is also a good opportunity to write your response paper since you will have read the material in more detail.)


**Research Design paper (30% total)**: Students will propose a research design to answer a question of interest. This paper must include an introduction, a literature review that situates the present question in the broader field, an answer to the question (the argument/theory), and a proposed design to answer the question. Students should meet with the instructor to discuss the paper and think about how one might design a study to answer the question of interest. This paper will be broken down into four parts that will be due throughout the semester.

1. Research question and list of literature (5% of final grade)
2. Literature review (5% of final grade)
3. Theory and research design (5% of final grade)
4. Full research design paper (15% of final grade)

- Papers must be submitted via Canvas. Late submissions will be penalized: half a point (out of 10) if turned in late but within 24 hours and one point for each 24-hour period after the due date.



## Course Conduct

In this class, we will discuss how citizens engage in politics and we will do so in a respectful environment. As is bound to be the case when we talk about politics, we may discuss topics that are controversial or of a sensitive nature for some or all in our group. Our goal in this class is to have constructive discussions about politics by understand how our psychology shapes political engagement. Derogatory comments toward other students or members of a particular social group will not be tolerated.


## Communication Policy

If you have questions about class in general or about a specific assignment, you can email me or come by my office. You can expect a response within 24 hours (unless you email me on Friday). Students are also more than welcome to attend office hours, as this is time set aside specifically for you. In addition, if my office door is open, feel free to stop by. Please note that I will not be "pre-grading" assignments, but I am happy to answer clarifying questions about them either via email or, preferably, during office hours.


## Regrading Policy

If a student wishes to appeal a grade, they must do so in writing within one week of receiving their graded assignment back. The student must indicate clearly what part of the assignment's grading they disagree with, making sure to respond to the instructor's comments in that section and making their case for why they should not have been deducted points. These requests should be done via email and the subject name should read "POL 310 Regrading request - [Name of assignment]." 


## Honor Code

Students are expected to do their own work and cite sources appropriately. Plagiarism will not be tolerated. In other words, students will be held to the standards of the Agnes Scott Honor System to which they agreed when enrolling at this institution. The Agnes Scott College honor code embodies an ideal of character, conduct, and citizenship, and is an important part of the College’s mission and core identity. This applies especially to academic honesty and integrity. Passing off someone else’s work (or Chat GPT's) as your own represents intellectual fraud and theft and violates the core values of our academic community. To be honorable, you should understand not only what counts as academic dishonesty, but also how to avoid engaging in these practices.


You are responsible for the content of any work submitted for this course. The use of artificial intelligence (AI) to generate a first draft of text is permitted, but you must review and revise any AI-generated text before submission. AI text generators can be useful tools but they are often prone to factual errors, incorrect or fabricated citations, and misinterpretations of abstract concepts. I neither encourage nor discourage their use, but utilize them with caution.



## Course Accessibility and Academic Accommodations

Agnes Scott College views disabilities as an integral part of the rich diversity of our community and strives to make all learning experiences as accessible as possible.  If you are a student who receives academic accommodations through the Office of Accessible Education, please schedule a meeting with me within the first two weeks of classes to discuss how your accommodations will be implemented for this course.  During this meeting, you are not expected to disclose any details concerning your disability, though you may discuss these details at your discretion.

If you are a student with a disability---physical, medical, psychological, or learning-specific---and have not connected with Accessible Education to discuss your accessibility needs, please visit the [main Office of Accessible Education webpage](https://www.agnesscott.edu/academicadvising/accessible-education/index.html) to learn more about accommodations, helpful resources and support, available through the Office of Accessible Education.  Students who register for accommodations during the semester should schedule a meeting with me after accommodations have been approved by the Office of Accessible Education.


## Wellbeing and Mental Health

The wellbeing and mental health of students is important; if you are having trouble completing your coursework, please reach out to the [Wellness Center](https://www.agnesscott.edu/wellnesscenter/index.html). Agnes Scott College provides cost-free mental health services to help you manage personal challenges that threaten your personal or academic well-being. If you believe you are experiencing unusual amounts of stress, sadness, or anxiety, please contact the Wellness Center and ask about their [Counseling and Psychological Services](https://www.agnesscott.edu/wellnesscenter/counseling/index.html).


## Title IX

Agnes Scott is here to help you if you have experienced any form of sexual harassment or violence, dating or domestic violence, or stalking. Please talk to any faculty or staff member with whom you feel comfortable. Faculty and staff members want to support you and have been trained to help. They will also inform the Title IX office so that you learn about options available to you. If you do not want college administrators to know what you have experienced, you may talk to the chaplain, as well as nurses or counselors in the Wellness Center with complete confidentiality. They will not tell anyone what you share with them unless you give your express permission. You may contact the Title IX Coordinator directly at <T9Coordinator@agnesscott.edu>.


## Diversity and Inclusion

Agnes Scott is a diverse and inclusive community. As one of the most diverse colleges in the nation, ASC is ideally positioned to be the model of a diverse and inclusive community that society can aspire to be. Such diversity raises the intellectual quality of the classroom experience, creating a unique environment for learning to understand and navigate the challenges of our times. By studying, living, and playing together, Agnes Scott College’s remarkably diverse student body hones the habits of mind, skills, and knowledge essential to ethical and innovative leadership in our increasingly heterogeneous and global society.  As such, this course adheres to the principles of diversity and inclusion as integral to the Agnes Scott community and respects people from all backgrounds. As a first step, this course affirms people’s decisions about gender expression and identity and will use each other’s preferred names and gender pronouns at all times.


## Center for Writing and Speaking

The Center for Writing and Speaking offers one-on-one appointments with trained peer tutors who can help you improve your written and oral communication skills. You are encouraged to use this resource when working on your class assignments to make your arguments more compelling, your sentences clearer, and persistent grammatical errors nonexistent. You can schedule an appointment at <https://www.agnesscott.edu/center-for-writing-and-speaking/index.html>.


## Syllabus Change Policy

This syllabus is only a guide for the course and is subject to change with advanced notice.


::: {.content-visible when-format="pdf"}
  
## Schedule

{{< include content/01-week1.qmd >}}

{{< include content/02-week2.qmd >}}

{{< include content/03-week3.qmd >}}

{{< include content/04-week4.qmd >}}

{{< include content/05-week5.qmd >}}

{{< include content/06-week6.qmd >}}

{{< include content/07-week7.qmd >}}

{{< include content/08-week8.qmd >}}

{{< include content/09-week9.qmd >}}

{{< include content/10-week10.qmd >}}

{{< include content/11-week11.qmd >}}

{{< include content/12-week12.qmd >}}

{{< include content/13-week13.qmd >}}

{{< include content/14-week14.qmd >}}

{{< include content/15-week15.qmd >}}

{{< include content/16-week16.qmd >}}

{{< include content/17-week17.qmd >}}

:::

{{< include index_title.qmd >}}
